---
title: "R Notebook"
output: html_notebook
---


# LC Summer Coding Club
## Session 1: Loading Data in RStudio Server

Welcome to Summer Coding Club part I! Today we're going to look at loading datasets into RStudio Server. There are lots of options, so let's get started:



* Built-in and package-specific datasets
* Uploading to the Files area
* From remote URL
* From Google Sheets



### Built-in datasets.
R has a number of built-in data sets, many of which are great for gaining some practice with getting your feet wet with R. 


```{r}

# This command will show the available built-in data sets currently loaded with "base R"
data()



```

We can learn more about a specific dataset by using the "?" command. We can also take a look at the data by using the "head" command on the dataset, or just priting the dataset:


```{r}
# Displays more info about "mtcars" in the help window
?mtcars

# Displays structure and first 6 rows
head(mtcars)

#Displays entire dataset
mtcars


```


Datasets can also be included with some libraries/packages. Let's load the tidyverse, and then run the data() command again:

```{r}
# The "tidyverse" is a series of packages used frequently in data science
library(tidyverse)

data()



```

YOUR TURN:
Load the "palmerpenguins" package, and take a look at the included dataset(s). How might you find out what they are?


```{r}



```



### Loading data from the files area

Each user on our RStudio Server instance has a home directory, where you can keep various files, R Projects, etc. You can view your files in the lower right-hand window, under the "Files" tab.

You should see a file called "multco.csv", which has some raw population numbers for Multnomah County over the last several years (data from data.census.gov). Below, you can use the "read_csv" function to open a dataset, and assign it to a variable (in this case, "multco"):



```{r}

# The "read_csv" function is part of the "readr" package, which is part of the tidyverse.
multco<-read_csv("multco.csv")

multco
```


You can also use the "upload" button in the files area to load a csv file, and then use that data with the technique above.

YOUR TURN:

Find a dataset in csv format, download it, upload it into your "files" area, and then assign it to a variable and print it out. Feel free to search the web for a dataset, or use EJ's Political Science data resource page: https://library.lclark.edu/polisci/ps_data

```{r}





```





### Loading data by URL

Many data sources provide stable links to comma separated value (csv) datasets. While you can use the download/upload method, you can also query a URL directly. This could be useful in cases where the data is dynamic.

```{r}

# The New York Times posts their updated daily covid data on their GitHub site (https://github.com/nytimes/covid-19-data). Below we'll load the data tracking statistics by US State:


url<-"https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv"


# The read_csv function also works for URLs
covidData<-read_csv(url)

covidData




```


YOUR TURN:

Peruse the datasets on FiveThirtyEight's github site: https://github.com/fivethirtyeight/data

Find a dataset you're interested in, and load it as a variable into RStudio Server.
(Hint: once you drill down to a csv file you want to load, click "raw" to get a direct URL)
```{r}




```





### Loading data from Google Spreadsheets

While one option for getting Google sheets data is to download the sheet as CSV, and then upload it to RStudio Server, we can also use the 'googlesheets4' package to directly import it from Google Drive. This could be useful when using a dynamic data set...you can just re-run the querying code to get the latest data.

Before doing an example, let's collect some data using a Google Form:

https://forms.gle/nLMPv6WnyXF5cm539

The results of form are logged in this Google Sheet: 
https://docs.google.com/spreadsheets/d/1JNaKXWEythLlr0FJR8ePtcyFANyZ5BKM-3EW-GsYABg/edit?usp=sharing


Now, there's a little bit of weirdness with loading the Google Spreadsheet into RStudio. Once you run the code below, you'll need to interact with the console window in order to satisfy Google's authentication process (thanks for Kristen Bott at Reed for the info below):

* In your console, you will be asked if it is okay to “cache OAuth access credentials in the folder” followed by a URL. (This means: “Is it okay for R to gain access to your Google Sheets in the future, or would you like to only grant one-time access?”) Type “1” in the console for “yes”, agreeing to both one-time and future access, and click enter.

* Next, R will open a window in your browser asking for tidyverse access to your Google account. This will allow R to access your documents in the future when you want to read in files. Follow the directions in the browser and grant access.

* Finally, you will be asked to copy a URL from your browser window into your console, where you will paste it in as the “authorization code” and then click enter.


```{r}

library(googlesheets4)
options(gargle_oob_default = TRUE)



classData <- read_sheet("https://docs.google.com/spreadsheets/d/1JNaKXWEythLlr0FJR8ePtcyFANyZ5BKM-3EW-GsYABg/edit?usp=sharing")


classData

```



### Some super light statistics

To wrap things up, we're going to determine the mean and standard deviations of a couple columns in our dataset. To do this, we need to first create a variable that has all the values for a specific column. You can do this by using the "dataset$column" syntax:


```{r}

distance<-classData$`How many miles are you currently from campus?`

distance


```

Now we can pretty easily calculate the mean and standard deviation for the array of responses for 'distance':

```{r}

mean(distance)

sd(distance)

```


YOUR TURN:

Calculate the mean and standard deviations for the responses to the 'pets' question, and the 'height' question.


```{r}


```


Bonus: What is the Pearson Correlation coefficient between height and number of pets owned?
(http://www.sthda.com/english/wiki/correlation-test-between-two-variables-in-r)

```{r}
# requires ggpubr library
library("ggpubr")
```


Next week: data wrangling & penguins with dplyr!



